---
title: "Tidyverse Training"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# setup environment
To setup the environment, I would like to set the working directory and load in the libraries needed

```{r, echo=FALSE, message=FALSE,warning=FALSE}
#setup working directory
setwd("C:/Users/natha/Dropbox/R_training")

#load libraries
library(tidyverse)
library(forcats)

#alternatives to setwd()
library(here)
read_csv(here("data","mtcars.csv"))
# https://www.youtube.com/watch?v=QYrdsjBvZN4

```

# readr - import data
Main type of table is a called a tibble
Packages to use:
- readr package -- flat files 
- readxl -- xls 
- dbi and backend -- SQL 
- haven package -- SPSS, SAS, STATA 
- jsonlite -- json

```{r}
# readr
read_csv()
read_tsv()
read_fwf()
read_log()

# load a dataset
df <- read_csv("housing.csv")


# readxl
library(readxl)
read_excel(path = ".data/mpg.xlsx", sheet = "Sheet 1")

# data.table
library(data.table)
df.f <- fread(file = "./data/file.csv", sep = ",")

# database connection
library(RODBC)
library(DBI)
#setup connection
connection <- odbcConnect("sqlconnect")
datafromsql <- sqlQuery(connection, "select * from data")

data <- read.csv('test.csv', header = TRUE)
sqlSave(connection, data, tablename = "test", rownames = FALSE, append = TRUE)

-------------------------
library(RODBC)

myConn <- odbcDriverConnect(connection= "Driver = {SQL Server Native Client 11.0};server=localhost, database=movies, trusted_connection=yes;")

data = sqlFetch(myConn, "tblActor",colnames = FALSE, rownames = FALSE)

SQLcommand <- sqlQuery(myConn,
"SELECT * 
  FROM tblActors
"

```
# dplyr - data manipulation

## select()
Select columns to return

```{r}
# select

df %>%
  select(total_bedrooms, total_rooms, median_house_value)

# rename a column
renamed_df <- df %>%
  select(total_rooms:median_house_value, rooms = total_rooms)

# exclude something
exclude_df <- df %>%
  select(-longitude, -latitude)

# starts with m
startswithdf <- df %>%
  select(starts_with(match = "total"))

# select by location instead of name
df %>%
  select (1:2)

#select second to last column
df %>%
  select((ncol(df)-2):ncol(df))

# other commands
# select (ends_with("rooms"))
# select(starts_with(c("rooms", "bedrooms")))
# select(contains("room"))
# regular expression using matches
# select(matches("[pt]al"))
# select(num_range("wk", 10:15))
```

## seperate() 
Pulls apart one column into multiple columns

## unite()
Inverse of separate - it combines multiple columns into a single column

```{r}

separate_df <- df %>%
  separate(ocean_proximity, into = c("part1", "part2"), sep = " ", convert = TRUE)

# default is for underscores or specify a sep
table5 %>% 
  unite(new, century, year)

```

## pull() 
Take a single column and turn into a vector

```{r}
df %>%
  pull(hwy)

```

## mutate()
Create, modify, and delete column

##mutate_all()
Apple function over every column

## mutate_at()
Apply function over specific columns

## mutate_if()
Apply function over columns that meet a condition

```{r}

# calculate new column
df %>%
  mutate (affordability = median_house_value / (median_income*10000))

# remove variable
df %>%
  mutate(longitude = NULL)

# grouped mutate act at grouping
starwars %>%
 select(name, mass, homeworld) %>%
 group_by(homeworld) %>%
 mutate(rank = min_rank(desc(mass)))

df %>%
  mutate_at(.tbl = .,
            vars = c("month", "dayofmonth"),
            .funs = str_pad, 2, "left","0")
df %>%
   mutate_at(.vars = vars(y, z),
             .funs = list(~ ifelse(x, ., NA)))
df %>%
   mutate(across(c(y, z), ~ ifelse(x, ., NA)))

```


##pivot_longer() and pivot_wider()
```{r}

# tidyr functions that replace gather and spread

# tableA looks likes this
#>   country     `1999` `2000`
#> * <chr>        <int>  <int>
#> 1 Afghanistan    745   2666
#> 2 Brazil       37737  80488
#> 3 China       212258 213766

tableA %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")

#>   country     year   cases
#>   <chr>       <chr>  <int>
#> 1 Afghanistan 1999     745
#> 2 Afghanistan 2000    2666
#> 3 Brazil      1999   37737
#> 4 Brazil      2000   80488
#> 5 China       1999  212258
#> 6 China       2000  213766

#------------------------------

# tableB looks likes this
#>   country      year type           count
#>   <chr>       <int> <chr>          <int>
#> 1 Afghanistan  1999 cases            745
#> 2 Afghanistan  1999 population  19987071
#> 3 Afghanistan  2000 cases           2666
#> 4 Afghanistan  2000 population  20595360
#> 5 Brazil       1999 cases          37737
#> 6 Brazil       1999 population 172006362

tableB %>%
    pivot_wider(names_from = type, values_from = count)

#> # A tibble: 6 x 4
#>   country      year  cases population
#>   <chr>       <int>  <int>      <int>
#> 1 Afghanistan  1999    745   19987071
#> 2 Afghanistan  2000   2666   20595360
#> 3 Brazil       1999  37737  172006362
#> 4 Brazil       2000  80488  174504898
#> 5 China        1999 212258 1272915272
#> 6 China        2000 213766 1280428583

```

## filter() 

```{r}

df %>%
  filter (weeks_popular >= 20, artist == 'Drake')

```


## rename()

```{r}
df %>%
  rename(newname = oldname)

df %>%
  rename(mnfc = manufacturer, 
         mode = model)  

df %>%
  select(mnfc = manufacturer, everything())

```

## arrange
Used to sort rows
Default is ascending 

```{r}

df %>%
  arrange(population)

df %>%
  arrange(desc(population))

```

## distinct() 
Removes duplicate rows - similar to SQL

```{r}

df %>%
  distinct(song)

```

## group_by()
Use to aggregate rows and summarise values
Use ungroup() to remove grouping variable
Can group() and ungroup() multiple times

```{r}

df <- df %>%
  group_by (manufacturer, model) %>%
  mutate('mean hwy' = mean(hwy)) %>%
  ungroup()

```

## summarise ()
first, last, nth, n, n_distinct, IQR, min, max, mean, median, var, sd

## across()
Apply a function (or functions) across multiple columns

```{r}
df %>% 
  group_by(ocean_proximity) %>%
  summarise(hsh= mean(households))

df %>%
  group_by (song) %>%
  summarise (total_weeks_popuular = max(weeks_popular))

starwars %>%
  summarize(across(height:mass, mean, na.rm = TRUE))

starwars %>%
  summarize(across(where(is.numeric), mean, na.rm = TRUE))

starwars %>%
  summarize(across(where(is.numeric), list(min = min, max = max), na.rm = TRUE))

starwars %>% 
  group_by(species) %>% 
  filter(n() > 1) %>% 
  summarise(across(c(sex, gender, homeworld), 
                   ~ length(unique(.x))),
                   n=n())

#change to {function}_{column} for naming
starwars %>% 
  summarise(across(where(is.numeric),
                   .fns = list(median = median, mean = mean), 
                    na.rm = TRUE,
                   .names = "{.fn}_{.col}"))  


```

## summarize_at()

## summarize_if

```{r}
df %>%
  summarize_at(vars(lifeExp, pop), funs(mean, sd))

df %>%
  summarize_if(is.numeric, funs(mean, sd))
```

## window functions
Calculations done within the grouping

```{r}

df %>%
  group_by(country) %>%
  mutate (lap_pop = lag(pop, order_by = year),
          pop_chg = pop-lag_pop)

df %>%
  group_by(manufacturer) %>%
  mutate('car_id_1' = row_number()) %>%
  ungroup()

df %>% 
  mutate ('delta_distance' = distance - lag (distance, 1))

```

## case_when()

```{r}

# case_when used during mutate
starwars %>%
  select(name:mass, gender, species) %>%
  mutate(
    height_cat = case_when(
      height > 191 ~ "tall",
      height < 167 ~ "short",
      TRUE  ~ "average"
    )
  )

x <- 1:16

case_when(
  x < 5 ~ "less than 5",
  x < 10 ~ "less than 10",
  TRUE ~ as.character(x)
)

case_when(str_detect(string = fruit, pattern = "a") ~ "contains a",
      T ~ "does not contain a"))
```

## rownames_to_columns()
Change row names into a column

```{r}
head(state.x77)
#            Population Income Illiteracy Life Exp Murder HS Grad Frost   Area
# Alabama          3615   3624        2.1    69.05   15.1    41.3    20  50708
# Alaska            365   6315        1.5    69.31   11.3    66.7   152 566432

rownames_to_column(as.data.frame(state.x77), "State") %>% 
  str() %>% 
  as_tibble()

# 'data.frame':	50 obs. of  9 variables:
#  $ State     : chr  "Alabama" "Alaska" "Arizona" "Arkansas" ...
#  $ Population: num  3615 365 2212 2110 21198 ...

```


## rowwise()
Do functions by row versus across row (e.g., mean of several columns by row)
Something that is applied to a dataframe, similar to grouping

```{r}
# mean across all values in all rows
fruits %>% 
  mutate(height_mean = mean(c(height_1, height_2, height_3))) 

## # A tibble: 3 x 7
##   fruit      height_1 height_2 height_3 width weight height_mean
##   <chr>         <dbl>    <dbl>    <dbl> <dbl>  <dbl>       <dbl>
## 1 Banana            4      4.2      3.5     1   0.5         7.74
## 2 Strawberry        1      0.9      1.2     1   0.25        7.74

# mean across all values in each row
fruits %>% 
  rowwise(fruit) %>% 
  mutate(height_mean = mean(c(height_1, height_2, height_3)))

## # A tibble: 3 x 7
## # Rowwise:  fruit
##   fruit      height_1 height_2 height_3 width weight height_mean
##   <chr>         <dbl>    <dbl>    <dbl> <dbl>  <dbl>       <dbl>
## 1 Banana            4      4.2      3.5     1   0.5         3.9 
## 2 Strawberry        1      0.9      1.2     1   0.25        1.03



# You could combine c_across with column-wise across to compute the proportion of the total for each column:

rf %>% #our row-wise data frame
  mutate(total = sum(c_across(w:z))) %>% #total each row
  ungroup() %>% # ungroup the rows
  mutate(across(w:z, ~ .x / total)) # the .x represents each column

## # A tibble: 6 x 6
##      id     w     x     y     z total
##   <int> <dbl> <dbl> <dbl> <dbl> <int>
## 1     1 0.1   0.2   0.3   0.4     100
## 2     2 0.106 0.202 0.298 0.394   104
```


# stringr - working with strings

Package used to work with string data
Can use single '' and double quotes " "
Do not mix them

## str_detect()
Returns logical vector if pattern is found
Similar to base R grepl()

```{r}
str_detect(fruit,"a")

#exclude
str_detect(fruit,"a", negate = T)

# case statement
df %>%
  mutate (flag = case_when(str_detect(string = fruit, pattern = "a") ~ "contains a",
      T ~ "does not contain a"))

df %>%
  mutate ('counta' = str_detect(string = fruit, pattern = "a"))

```

## str_which()
Locates the indexes of strings which contain pattern match
Location of the match in a string

## str_count()
Counts the number of patterns in a given string; returns a vector of counts

```{r}
# returns a list of elements within vector that contains the pattern
str_which(string = fruit, pattern = "a")

#returns a tibble with fruit name and count
df %>%
  mutate ('counta' = str_count(string = fruit, pattern = "a"))

```

## str_locate()
Locates the position of the first match

## str_locate_all()

Locates the position of all matches

## str_pad()
add values if missing

## str_trunc()
Truncates data to a specific width

## str_trim
Removes characters

```{r}
# returns of start and end occurrences 
df %>%
  str_locate(fruit, pattern = "a") %>%
  as_tibble() %>%
  mutate(fruit=fruit) %>%
  select(fruit, start, end)

df %>%
  str_locate(fruit, pattern = "a")

#retuns list of vectors that match
df %>%
  str_locate_all(fruit, pattern = "a")

# adds pad to meet width requirements, can be left, right, or both
df %>%
mutate (month = str_pad(month, width = 2, side="left", pad = "0"))

str_pad(fruit, width = 20, side = "left", pad = "X")

# truncates and adds ellipsis for anything that is larger than width
str_trunc(fruit, width = 5, side = "left", ellipsis = "...")

# 
str_trim(string, side = "left")

```

## str_subset()  
## str_sub()
Returns only rows that contain pattern

```{r}

str_detect(string = fruit, pattern = "a")

# extract first three letters
str_sub(fruit, start = 1, end = 3)

# extract first letter and count word freq
words.df %>%
  mutate('first letter' = str_sub(word, 1,1))%>%
  count('first letter') %>%
  arrange(desc(n))

str_subset(fruit, pattern = 'c')

df %>%
  mutate(abbreviation = str_sub(country, 1, 3)
  
# return from end
df %>%
  mutate(flight_code = str_sub(flight_name, -3, -1)

# rereaokce
str_sub(string, start, end) <- "new string"     

# replace first matched case
str_replace(string, pattern, replacement)

# replace all matched cases
str_replace_all(string, pattern, replacement)
            
```

## str_replace()
Replace first matched case

## str_replace_all()
Replace all matched cases

## str_to_lower()
Replaces letters to lower case

## str_to_upper()
Replaces letters to upper case

## str_to_title()
Replaces letters to title case

```{r}
# replaces first three letters 
str_sub(df$fruit, 1, 3) <- "FRU"

# reaplces first occurence of a with A
str_replace(fruit, pattern = "a", replacement = "A")

str_to_lower(string)

str_to_upper(string)

str_to_title(string)

```

## str_c
Join multiple strings into a single string
String concatenate 
Collapse a vector strings into single string

## str_duup
Repeat a string multiple times

## str_split_fixed()
Split a vectoor strings into matrix of substrings

## str_glue()
Merge together strings and expressions

## str_glue_data()

```{r}

str_c(fruit1, fruit2, fruit3, sep = ", ")

#repeats the word multiple times
str_dup(fruit1,5)

# split based on whitespace
str_split_fixed (fruit, pattern = " ", n = 2)

# splits vector of strings into a list/matrix
str_split(sentences, pattern = " ")

name <- "nate"
str_glue("my name is {name}")

# use df to create strings
str_glue_data(mtcars, "The care {rownames(mtcars)}: {hp} horsepower, {cyl} number of cylinders")

```

## str_extract
Returns first match

## str_extract_all
Returns every match

## str_match()
Returns first match

## str_match_all()
Returns every match
Max count of pattern to create number of columns

```{r}
# returns vector  
str_extract(string = fruit, pattern = "c")

# returns list of occurences
str_extract_all(string = fruit, pattern = "c")

```

## str_length()
Returns the number of characters

```{r}
# 
str_length (string = fruit)

# fruits w ith 1o or more characteros
fruit[str_length (string = fruit) >= 10]

```

## str_view()
HTML rendering function - first regex match

## str_view_all()
HTML rendering function - all regex match

```{r}

str_view(string = fruit, pattern = "a")

str_view_all(string = fruit, pattern = "a", match = T)

```

# regular expressions 
Describing patterns in strings 
special characters or metacharacters
classes
alternates, anchors, and groups
look arounds and quantifiers

special characters 
.   match any characters
\   escape special character
()  define groups
[]  define classes / (one of)
{}  define quantifiers
^   match start of string
$ match end of a string
| OR operator
?   zero or one quantifier
*   zero or more quantifier 
+   1 or more quantifier 

Classes
[:digit:] digits
[:alpha:] letters 
[:lower:] lowercase
[:upper:] uppercase


Alternate
|     or
[]    one of
[^ ]  anything but
[ - ] range


(ab|cb) group syntax

Look arounds
(?=)    followed by       a(?=c)      after a, c    
(?! )   not followed by   a(?!c)      after a, NOT c
(?<= )  preceded by       (?<=b)a     before a, b
(?<! )  not preceded by   (?<!b)a     before a, NOT b


Quantifiers
?     zero or more      a? 
*     zero or more      a*
+     one or more       a+
{n}   exactly n         a{2}
{n,}  n or more         a{2,}
{n,m} between n and m   a{2,3}


```{r}

# find strings with digits
str_subset(string, pattern = "\\d")

# find strings without digits (at least one shown)
str_subset(string, "\\D")

# find strings matching a pattern (digit-digit-digit)
str_subset(string, "\\d-\\d-\\d")

# find strings with whitespace
str_subset(string, "\\s")

# ANCHORS
# find words that start with a
str_subset(words, "^a")

# find words that end with a
str_subset(words, "a$")

# find a specific word
str_subset(words, "^actor$")

# ALTERNATES
# find words that start with af or ag
str_subset(words, "^af|^ag")

# find words containing to letters x,y,or x
str_subset(words, "[xyz]")

# find words not containing certain letters (a to y) ^ means to not include, [] means one or more
str_subset(words, "[^[a-y]]")

# find all sentences that include words the or an
# group 1 = the
# group 2 = an
str_subset(sentences,"(\\sthe\\s|\\s\\san\\s)"

# find words with repeated pairs (back reference) - numbers reference to groups ()
str_subset(words, "(..)\\1")   # a groups of two letters needs to be repeated (e.g., isis)
str_subset(words, "(a)(b)") aba
str_subset(words, "(a)(b)\\1") aba
str_subset(words, "(a)(b)\\1\\2") abab


# look for words where w follows a
str_subset(words, "w(?=a)")

# look for words where w is not not followed by a
str_subset(words, "w(?!a)")


mutate(setences = str_remove_all(setences, "[:punct:]"), # remove punctuatin
       sentence = str_to_lower(sentence) %>% # make lowercase
       mutate('num words' = str_count(string = sentence, pattern = "\\s+")+1) # count number of spaces and add one

```


# forcats - working with factors

## fct_count ()
Returns the counts for each level

## fct_unique
Returns unique set of levels in a factor

```{r}

# return counts
df %>%
  .$manufacturer %>%
  fct_count()

# alternative
df %>%
  count(manufacturer)

# return unique list of factors
df %>%
  .$manufacturer %>%
  fct_unique()
```


## fct_c()
Combine factors with different levels

Reorder factor levels by first appearance, frequency, or numeric order

## fct_relevel()
Manually reorder factor levels

## fct_infreq()
by number of observations with each level (largest first)

## fct_inorder()
by the order in which they first appear

## fct_rev()
reverse order factors

## fct_inseq()
by numeric value of level

## fct_reorder()
(.f, .x, .fun = median,..., .desc = FALSE) 
Reorder levels by their relationship with another variable

## fct_reorder2()
(.f, .x, .y, .fun = last2, ..., .desc = TRUE) 
Reorder levels by their final values when plotted with two other variables


```{r}
# merge all levels together
f1 <- factor(c("a", "c"))
f2 <- factor(c("b", "a"))
fct_c(f1, f2)

# Relevel to the end
fct_relevel(f, "a", after = Inf)

# Move to the third position
fct_relevel(f, "a", after = 2)

# using a list
fct_relevel(f, c("b", "c", "a"))

# by order in data
fct_inorder(f)


boxplot(data = PlantGrowth, weight ~ reorder(group, weight))

ggplot(diamonds,aes(carat, price, color = fct_reorder2(color, carat, price))) + geom_smooth()



# number of observations (low to high)
df %>%
  mutate(manufacturer = fct_infreq(manufacturer))

# number of observations (low to high)
df %>%
  mutate(manufacturer = fct_infreq(manufacturer), 
         manufacturer = fct_rev(manufacturer))

# by numerical mapping of levels
fct_inseq(f)

```



## fct_lump()
A family for lumping together levels that meet some criteria.

## fct_lump_min()
lumps levels that appear fewer than min times

## fct_lump_prop()
lumps levels that appear in fewer prop * n times

## fct_lump_n() 
lumps all levels except for the n most frequent (or least frequent if n < 0)

## fct_lump_lowfreq() 
lumps together the least frequent levels, ensuring that "other" is still the smallest level


```{r}
starwars %>%
  filter(!is.na(species)) %>%
  mutate(species = fct_lump_n(species, n = 3)) %>%
  count(species)
#> # A tibble: 4 × 2
#>   species     n
#>   <fct>   <int>
#> 1 Droid       6
#> 2 Gungan      3
#> 3 Human      35
#> 4 Other      39

fct_lump_min(f, min, other_level = "Other")

```

## fct_recode()
Manually change levels

## fct_collapse()
Collapse groups into manually defined groups

## fct_other()
Replace levels with other

## fct_drop()
Drop unused levels

## fct_expand()
Add levels to a factor

## fct_explicit_na()
Missing data will be put into level x

```{r}
gss_cat %>% 
  mutate(partyid = fct_recode(partyid, 
    "Republican strong" = "Strong republican", 
    "Republican weak" = "Not str republican",
    "Independent near rep" = "Ind near rep")) %>% 
  count(partyid)

gss_cat %>% 
  mutate(partyid = fct_collapse(partyid, 
    other = c("No answer", "Don't know", "Other party"), 
    rep = c("Strong republican", "Not str republican"), 
    ind = c("Ind,near rep", "Independent", "Ind,near dem"), 
    dem = c("Not str democrat", "Strong democrat") )) %>%
count(partyid)

# collapses to Other anything with less than 2
fct_lump_min(f, min = 2)

fct_other(f, keep = c("a", "b"))

# drops factor x
f5 <- factor(c("a","b"),c("a","b","x"))
f6 <- fct_drop(f5)

# adds a factor x
fct_expand(f6, "x")

# NA assigned a level so they are shown in plots
fct_explicit_na(factor(c("a", "b", NA)))

```






























# Logical Conditions and Loops

## if() and else{}
if(condition) {
}

## else if()

## ifelse()
Returns TRUE, FALSE, or NA

## while loop


## for loop



```{r}
x <- 5

if(x < 0) {
  print("x is a negative number")
} else if (x == 0){
  print("x is 0")
} else {
  print("x is a positive number")
}

ifelse(example > 0, "Positive", "Not Positive")

#ifelse within mutate
df %>% 
  mutate (short_country =
            ifelse(country == "Bosnia", "B", as.character(country))
          )

# statement to test, positive response, negative response
ifelse(4 > 5, "Yes!", "No")


# example of a while loop
a = 5
b = 1

while(a > b){
  print("a is greater than b")
  a <- a + 1
}

# for every number if 1 to 5, the result will be 1 to power of 2
result <- c()

for (i in 1:5) {
  result[i]<- i ^ 2  
}

```





















## case_when()
Multiple conditions within multiple variables
Take care of NAs first


case_when (condition1 ~ rez1, 
condition2 ~ rez2
TRUE ~ rez_else)




```{r}
df %>%
  mutate(short_country = 
      case_when(
        country == "Bosnia" ~ "B",
        country == "Germany" ~ "G",
        TRUE ~ as.character(country)
                )     
         )





```


#row_number
Ranks given rows and works within group_by

df %>%
  mutate('car_id' = row_number())





# lubridate

Working with Dates and Times
lubridate package

```{r}
library(lubridate)

today()

now()

ymd("2017-01-31")
mdy("January 31st, 2017")
dmy("31-Jan-2017")
ymd(20170131)

ymd_hms("2017-01-31 20:11:59")
mdy_hm("01/31/2017 08:01")

make_date()

make_datetime()


```
Time spans
durations, which represent an exact number of seconds.
periods, which represent human units like weeks and months.
intervals, which represent a starting and ending point.

In R, when you subtract two dates, you get a difftime object:

```{r}
h_age <- today() - ymd(19791014)
h_age
as.duration(h_age)

dseconds(1)
dminutes(1)
dhours(1)
ddays(1)
dweeks(1)
dyear(1)


```

```{r}

```

# function(purrr)
R is a functional programming (FP) language
Iteration is completed using a for loop or applying a function across an object


map (apply a function, control the output, shortbuts, several arguments)

working with lists
nested data

```{r}

```



















#rlang



```{r}

```


# Joins


```{r}

planes %>% 
  count(tailnum) %>% 
  filter(n > 1)

flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")

x %>% 
  inner_join(y, by = "key")

```





```{r}
gss_cat %>% 
  mutate(relig = fct_lump(relig)) %>% 
  count(relig)

gss_cat %>%
  mutate(relig = fct_lump(relig, n = 10)) %>%
  count(relig, sort = TRUE) %>%
  print(n = Inf)

```










# purr
Functional programming tools 
map() applies iteratively to each element of a list or vector

map(.x, .f, ...)

map() - returns list
map_chr() - returns char vector
map_lgl() - returns integer vector
map_dbl() - returns double vector
map_dfc() - returns data frame (column bind) - join
map_dfr() - returns data frame (row bind) - stacked / union



```{r}

df %>%
  select(total_bedrooms, total_rooms) %>%
  map_dfr(.f = mean, na.rm = TRUE)

# example for sets of columns



# fit multiple lm models based on category

df %>%
  split(.$households) %>%
  map(function(df) lm(formula=hwy ~ disp, 
                      data = df))
#shorter version
df %>%
  split(.$households) %>%
  map(~lm(hwy ~ disp, data = .))

# extract r2 for each model

models %>%
  map(summary) %>%
  map_dbl("r.squared")

```

## walk()

```{r}
# create multiple plots

plots <- mpg %>%
  split(.$manufacturer) %>%
  map(~ggplot(.,aes(displ, hwy))+
      geom_point()+ 
      ggtitle(.$manufacturer)
      )



```



## pluck

## keep
select elements that pass a logical test

## discard
select elements that do not pass a logical test


create directory
export multiple files
directory <- ".data/mpg_export"

if(!dir.exists(directory)){
  dir.create(directory)
  }



```{r}

## create plots with map (list of plots)
plots <- mpg %>% 
  split(.$manufacturer) %>% 
  map(~ggplot(., aes(displ, hwy)) + geom_point() + ggtitle(paste0(.$manufacturer)))

## create directory where plots will be stored
if(!dir.exists("./data/pwalk_plots")){
  dir.create("./data/pwalk_plots")
}

## export figures
list(str_c("./data/pwalk_plots/",names(plots), ".pdf"),  # list of pairs: plot path & plot
     plots) %>% 
  pwalk(., ggsave) # save figures


## check if directory exists for storing the plots  
directory <- "./data/mpg_plot_export/" # path to directory

## create directory if neccessary
if(!dir.exists(directory)){
  dir.create(directory)
}

## add a path for each plot
df.plots <- df.plots %>% 
  mutate(plot_path = paste0(directory, manufacturer, ".png"))
  
## create a list of plot path and plot  for pmap to export plots
list(plot     = df.plots$plot,            # plot object
     filename = df.plots$plot_path) %>%   # path to where write .png files
  pmap(.l = ., .f = ggsave) %>%           # do the export
  quietly()  # suppress console output!!!


```






# Conditions in R

You can use || (or) and && (and) to combine multiple logical expressions.

if (this) {
  # do that
} else if (that) {
  # do something else
} else {
  # 
}

But if you end up with a very long series of chained if statements, you should consider rewriting. One useful technique is the switch() function. It allows you to evaluate selected code based on position or name.

function(x, y, op) {
   switch(op,
     plus = x + y,
     minus = x - y,
     times = x * y,
     divide = x / y,
     stop("Unknown op!")
   )
 }
